\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{amsmath}

\lstset{basicstyle=\ttfamily\footnotesize, breaklines=true, frame=single}

\title{Experiment Results Analysis:\\PagedKVCOMM Ablation Study (April 5, 2026)}
\author{}
\date{}

\begin{document}
\maketitle

%-----------------------------------------------------------------------
\section{Experiment Overview}
%-----------------------------------------------------------------------

Two jobs were submitted on April 5, 2026, targeting an RTX Pro 6000 GPU node
(\texttt{compsci-gpu} partition).  Table~\ref{tab:exp_overview} summarises
their configuration.

\begin{table}[h]
\centering
\caption{Experiment configuration summary.}
\label{tab:exp_overview}
\resizebox{\textwidth}{!}{%
\begin{tabular}{lll}
\toprule
\textbf{Property} & \textbf{Job 10939240 (Stage-1 Baseline)} & \textbf{Job 10939241 (HF CRS Ablation)} \\
\midrule
SLURM Job ID      & 10939240 & 10939241 \\
Script            & \texttt{KVCOMM/run\_ablation\_stage1.sh} & \texttt{scripts/run\_hf\_crs\_ablation.sh} \\
Log               & \texttt{KVCOMM/logs/4-5\_page\_abl\_stage1\_rtx\_pro/} & \texttt{logs/hf\_crs\_ablation\_rtx\_pro\_10939241.log} \\
                  & \quad\texttt{page\_abl\_stage1\_rtx\_pro\_10939240.log} & \\
Result dir        & \texttt{KVCOMM/result/ablation\_1\_baseline\_rtx\_pro/} & \texttt{KVCOMM/result/hf\_crs\_ablation\_crs\_rtx\_pro\_20260405\_131449/} \\
Backend           & \texttt{gpt\_chat} & HuggingFace (\texttt{hf}) \\
\texttt{KVCOMM\_PAGED}   & 0 (disabled) & -- \\
CRS               & \textbf{OFF} (\texttt{--no-current-round-sharing}) & \textbf{ON} \\
Cross-round KV reuse & ON (\texttt{--execution\_mode allow\_kv\_reuse}) & ON \\
Flash Attention   & OFF & -- \\
Completion        & \texttt{[1/3] Completed} (stage 1 of 3) & \texttt{[1/1] Done} \\
\bottomrule
\end{tabular}}
\end{table}

\noindent\textbf{Important note:} The two jobs use different inference backends
(\texttt{gpt\_chat} vs.\ HuggingFace) and differ in CRS status simultaneously.
They are \emph{not} a clean single-variable comparison; direct accuracy
comparisons should be interpreted with this caveat in mind.

%-----------------------------------------------------------------------
\section{Overall Accuracy}
%-----------------------------------------------------------------------

\begin{table}[h]
\centering
\caption{Final cumulative accuracy per agent node at end of run.}
\label{tab:overall_acc}
\begin{tabular}{lcccc}
\toprule
\textbf{Job} & \textbf{Overall} & \textbf{Node 1} & \textbf{Node 2} & \textbf{Node 3} \\
\midrule
10939240 (gpt\_chat, no CRS) & 0.8218 & 0.7726 & 0.3730 & 0.7445 \\
10939241 (hf, CRS ON)        & 0.7726 & 0.7748 & 0.3904 & 0.6588 \\
\midrule
$\Delta$ (CRS ON $-$ baseline) & $-0.049$ & $+0.002$ & $+0.017$ & $-0.086$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Observations.}
\begin{itemize}
  \item Overall accuracy is higher in the baseline (0.822 vs.\ 0.773), but the
        two jobs differ in both backend and CRS, so the gap cannot be attributed
        to CRS alone.
  \item Node~1 accuracy is nearly identical across jobs ($\approx$0.77),
        suggesting that the Math Solver role is robust to both backend and CRS
        changes.
  \item Node~2 is consistently the weakest node in both jobs ($\approx$0.37--0.39).
        See Section~\ref{sec:node2} for a detailed analysis.
  \item Node~3 drops noticeably when CRS is enabled (0.744 $\to$ 0.659,
        $\Delta = -0.086$), making it the primary driver of the overall
        accuracy gap.
\end{itemize}

%-----------------------------------------------------------------------
\section{Time to First Token (TTFT)}
%-----------------------------------------------------------------------

\begin{table}[h]
\centering
\caption{TTFT statistics broken down by execution mode, aggregated over all
         agent nodes.}
\label{tab:ttft}
\begin{tabular}{llcccc}
\toprule
\textbf{Job} & \textbf{Mode} & \textbf{N} & \textbf{Avg (s)} & \textbf{Min (s)} & \textbf{Max (s)} \\
\midrule
\multirow{2}{*}{10939240 (no CRS)} & dense\_prefill & 1573 & 0.0770 & 0.0439 & 0.1208 \\
                                   & kv\_reuse      & 3703 & 0.2740 & 0.0378 & 1.2847 \\
\midrule
\multirow{2}{*}{10939241 (CRS ON)} & dense\_prefill &  866 & 0.0792 & 0.0523 & 0.1242 \\
                                   & kv\_reuse      & 4410 & 0.2755 & 0.0373 & 1.3067 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Observations.}
\begin{itemize}
  \item TTFT is nearly identical between the two jobs for the same mode,
        indicating that the backend and CRS differences do not affect
        time-to-first-token.
  \item \texttt{kv\_reuse} TTFT ($\approx$274--276\,ms) is \textbf{$\approx$3.5$\times$
        higher} than \texttt{dense\_prefill} ($\approx$77--79\,ms).  This is
        counter-intuitive: reusing a cached KV state incurs more latency than a
        fresh prefill, suggesting that anchor retrieval and KV loading dominate
        the overhead.
  \item \texttt{kv\_reuse} max TTFT reaches 1.28--1.31\,s (17$\times$ higher
        than average), indicating high variance likely caused by cache miss
        penalties or large anchor KV tensors.
  \item CRS ON increases the number of \texttt{kv\_reuse} invocations
        (4410 vs.\ 3703), consistent with CRS providing more anchor candidates
        for matching.
\end{itemize}

%-----------------------------------------------------------------------
\section{Node 2 Analysis: Mathematical Analyst}
\label{sec:node2}
%-----------------------------------------------------------------------

\subsection{Prompt Design}

Node~2 (Mathematical Analyst) uses the following system prompt:

\begin{lstlisting}
You are a mathematical analyst.
You will be given a math problem, analysis and code from other agents.
You need to first analyze the problem-solving process step by step,
where the variables are represented by letters.
Then you substitute the values into the analysis process to perform
calculations and get the results.
The last line of your output contains only the final result without
any units, for example: The answer is 140
You will be given some examples you may refer to.
\end{lstlisting}

\noindent The few-shot exemplars for Node~2 follow a two-phase template format:

\begin{enumerate}
  \item \textbf{Abstract phase}: describe the solution using placeholder variable
        names (e.g., \texttt{\{ori\_tree\_num\}}, \texttt{\{after\_planted\_tree\_num\}}).
  \item \textbf{Instantiation phase}: substitute concrete values and compute the
        answer.
\end{enumerate}

Only \textbf{two} few-shot examples are provided (both involving very simple
addition/subtraction), compared to three richer examples for Node~1.

\subsection{Accuracy Breakdown by Execution Mode}

\begin{table}[h]
\centering
\caption{Node 2 accuracy broken down by execution mode.}
\label{tab:node2_mode}
\begin{tabular}{llcccc}
\toprule
\textbf{Job} & \textbf{Mode} & \textbf{Correct} & \textbf{Wrong} & \textbf{Total} & \textbf{Acc} \\
\midrule
\multirow{2}{*}{10939240 (no CRS)} & dense\_prefill & 209 & 176 & 385  & 0.5429 \\
                                   & kv\_reuse      & 283 & 651 & 934  & 0.3030 \\
\midrule
\multirow{2}{*}{10939241 (CRS ON)} & dense\_prefill &  57 &  62 & 119  & 0.4790 \\
                                   & kv\_reuse      & 458 & 742 & 1200 & 0.3817 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Observations.}
\begin{itemize}
  \item In the baseline, Node~2 accuracy drops from 0.54 (dense\_prefill) to
        0.30 (kv\_reuse), a \textbf{$-$0.24 degradation}.  With wrong $=$ 2.3$\times$
        correct under kv\_reuse, KV reuse is the primary driver of Node~2
        failures, not the prompt design alone.
  \item The template-style variable-placeholder format appears particularly
        \textbf{sensitive to KV context pollution}: when a cached KV from a
        different problem is reused, the model tends to substitute wrong variable
        values or copy the previous problem's abstract template, as observed in
        FinalRefer outputs (e.g., sheep problem answered with trees analysis).
  \item With CRS ON, kv\_reuse accuracy of Node~2 \emph{recovers} slightly
        (0.30 $\to$ 0.38), suggesting that cross-agent sharing within the same
        round provides a stabilising signal that partially counteracts KV
        pollution.
  \item dense\_prefill sample count drops sharply under CRS (385 $\to$ 119),
        since CRS causes more requests to satisfy anchor-matching conditions
        and fall into kv\_reuse.
\end{itemize}

%-----------------------------------------------------------------------
\section{Node 3 Analysis: Programming Expert}
\label{sec:node3}
%-----------------------------------------------------------------------

\subsection{Accuracy Breakdown by Execution Mode}

\begin{table}[h]
\centering
\caption{Node 3 accuracy broken down by execution mode.}
\label{tab:node3_mode}
\begin{tabular}{llcccc}
\toprule
\textbf{Job} & \textbf{Mode} & \textbf{Correct} & \textbf{Wrong} & \textbf{Total} & \textbf{Acc} \\
\midrule
\multirow{2}{*}{10939240 (no CRS)} & dense\_prefill & 280 & 125 & 405  & 0.6914 \\
                                   & kv\_reuse      & 702 & 212 & 914  & 0.7681 \\
\midrule
\multirow{2}{*}{10939241 (CRS ON)} & dense\_prefill & 101 &  60 & 161  & 0.6273 \\
                                   & kv\_reuse      & 768 & 390 & 1158 & 0.6632 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Observations.}
\begin{itemize}
  \item Node~3 exhibits the \textbf{opposite pattern to Node~2}: in the
        baseline, kv\_reuse accuracy (0.768) is \emph{higher} than
        dense\_prefill (0.691), a $+$0.077 gain.  Reusing KV from similar
        problems appears to provide useful structural context for Python code
        generation.
  \item The Programming Expert prompt enforces strict output format
        (a self-contained Python function), which likely makes it more
        robust to KV context from other problems---the code structure acts
        as a natural ``guard'' against template pollution.
  \item With CRS ON, both modes degrade: dense\_prefill $0.691 \to 0.627$,
        kv\_reuse $0.768 \to 0.663$.  Node~3 is the \textbf{main contributor}
        to the $-$0.086 overall accuracy drop under CRS, suggesting that
        the additional cross-agent context introduced by CRS is harmful for
        code generation (possibly introducing natural-language reasoning
        patterns that conflict with the strict code-only output constraint).
  \item This is consistent with the prompt instruction: \emph{``Do not include
        anything other than Python code blocks in your response.''} CRS may
        inject reasoning text from Node~1 or Node~2 into the KV context,
        violating this constraint implicitly.
\end{itemize}

%-----------------------------------------------------------------------
\section{Cross-Node and Cross-Mode Summary}
%-----------------------------------------------------------------------

\begin{table}[h]
\centering
\caption{Accuracy summary across all nodes and modes for both jobs.}
\label{tab:summary}
\begin{tabular}{llccc}
\toprule
\textbf{Job} & \textbf{Mode} & \textbf{Node 1} & \textbf{Node 2} & \textbf{Node 3} \\
\midrule
\multirow{2}{*}{10939240 (no CRS)}
  & dense\_prefill & -- & 0.5429 & 0.6914 \\
  & kv\_reuse      & -- & 0.3030 & 0.7681 \\
\midrule
\multirow{2}{*}{10939241 (CRS ON)}
  & dense\_prefill & -- & 0.4790 & 0.6273 \\
  & kv\_reuse      & -- & 0.3817 & 0.6632 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Key takeaways.}
\begin{itemize}
  \item \textbf{Node 2 (Mathematical Analyst)} is consistently the weakest
        node.  Its template-based prompt is highly vulnerable to KV context
        pollution; kv\_reuse degrades accuracy by $-$0.17 to $-$0.24.
  \item \textbf{Node 3 (Programming Expert)} benefits from kv\_reuse in
        isolation (no CRS), but is harmed by CRS.  Its strict code-only output
        format provides partial robustness to KV pollution, but cross-agent
        context from CRS conflicts with the code generation task.
  \item \textbf{kv\_reuse TTFT is 3.5$\times$ higher} than dense\_prefill,
        undermining the latency motivation for KV reuse.  The overhead likely
        originates from anchor retrieval and KV tensor loading, not from
        actual generation.
  \item The two experiments are not a clean ablation (different backends and
        CRS status co-vary).  A proper ablation should fix the backend and vary
        only CRS, or vary only paged vs.\ non-paged, to isolate each effect.
\end{itemize}

\end{document}
